---
title: "Point Pattern Analysis (Density_Based)"
author: "Afiq"
date: "2023-03-20"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
```{r}
library(readxl)
library(sf)
library(ggplot2)
library(rgdal)
library(here)
library(tmap)
library(dplyr)
library(spatstat)
```
# Load HFMD Data 
```{r}
HFMD_overall <- read_excel( here("HFMD_17_22.xlsx"))
HFMD_overall <- st_as_sf(HFMD_overall,coords = c("X_LONGITUDE","X_LATITUDE"), 
                    crs = 4326)
HFMD_overall <- st_transform(HFMD_overall, 3168)
Penang <- st_read(here("Penang", 
                  "mukim2000_penang1.shp"))
```
#View Data
```{r}
st_crs(Penang)<-3168
st_crs(Penang)
```
```{r}
glimpse(HFMD_overall)
```
# Plot HFMD points(2017-2022) in Penang Polygon
```{r}
tm_shape(Penang)+ tm_polygons("NAMA_DP", title = "HFMD Cases in Pulau Pinang 2017-2022",title.size = 10)+tm_shape(HFMD_overall) + tm_dots(col = "black")+ tm_layout(legend.outside = TRUE)
```

## Plot Yearly HFMD Cases in Pulau Pinang

```{r}
tm_shape(Penang) +
  tm_polygons("NAMA_DP", title = "HFMD Cases Between 2017-2022",title.size = 1) + 
  tm_shape(HFMD_overall) + tm_layout( legend.outside = TRUE) + tm_dots(col = "black", size = 0.03 ) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
```

## Plot HFMD points(2017-2022) according to district

```{r}
tm_shape(Penang[Penang$NAMA_DP == "TIMUR LAUT",])+
  tm_polygons("NAMMUK",title = "HFMD Cases in TIMUR LAUT 2017-2022")+
  tm_shape(HFMD_overall[HFMD_overall$DAERAH == "TIMUR LAUT",])+tm_dots (col = "black")+tm_layout(legend.outside = TRUE) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
tm_shape(Penang[Penang$NAMA_DP == "BARAT DAYA",])+
  tm_polygons("NAMMUK",title = "HFMD Cases in BARAT DAYA 2017-2022")+
  tm_shape(HFMD_overall[HFMD_overall$DAERAH == "BARAT DAYA",])+tm_dots (col = "black")+tm_layout(legend.outside = TRUE) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
tm_shape(Penang[Penang$NAMA_DP == "SP.TENGAH",])+
  tm_polygons("NAMMUK",title = "HFMD Cases in SP.Tengah 2017-2022")+
  tm_shape(HFMD_overall[HFMD_overall$DAERAH == "Seberang PERAI TENGAH",])+tm_dots (col = "black")+tm_layout(legend.outside = TRUE) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
tm_shape(Penang[Penang$NAMA_DP == "SP.UTARA",])+
  tm_polygons("NAMMUK",title = "HFMD Cases in SP.UTARA 2017-2022")+
  tm_shape(HFMD_overall[HFMD_overall$DAERAH == "SEBERANG PERAI UTARA",])+tm_dots (col = "black")+tm_layout(legend.outside = TRUE) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
tm_shape(Penang[Penang$NAMA_DP == "SP.SELATAN",])+
  tm_polygons("NAMMUK",title = "HFMD Cases in SP. Selatan 2017-2022")+
  tm_shape(HFMD_overall[HFMD_overall$DAERAH == "Seberang PERAI SELATAN",])+tm_dots (col = "black")+tm_layout(legend.outside = TRUE) + tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
```

## Plot Monthly HFMD Cases in Pulau Pinang

### 1.Convert "MONTH" variable to factor with ordered levels

```{r}
HFMD_overall$MONTH <- factor(HFMD_overall$MONTH, 
                             levels = month.name, 
                             ordered = TRUE)
```
#### Plotting monthly HFMD Map
```{r}
tm_shape(Penang) +
  tm_polygons("NAMA_DP", title = "Monthly HFMD Cases",title.size = 1) + 
  tm_shape(HFMD_overall) + tm_layout( legend.outside = TRUE) + tm_dots(col = "black", size = 0.03 ) + tm_facets("MONTH", nrow = 2, ncol = 6, as.layers = TRUE)
```
# Incidence of HFMD in Pulau Pinang 
# 1.Prepare Data
# 2.count point in each Mukim 
```{r}
Overall_Coor <- HFMD_overall [,c('geometry','YEAR','DAERAH')]
Overall_Coor <- na.omit(Overall_Coor)
```
```{r}
intersection_overall <- st_intersection(x=Penang,y=Overall_Coor)
count_HFMD_Mukim_Overall <- intersection_overall %>% count (YEAR,NAMA_DP,NAMMUK,PENDUDUK) 
```
# 3.Load population data
```{r}
Gen_pop <- read_excel(here("Gen_Pop.xlsx"))
```
# 4.calculate incidence
```{r}
HFMD_Penang_Overall <- merge(count_HFMD_Mukim_Overall,Gen_pop, by = c("YEAR", "NAMA_DP","NAMMUK") , all=TRUE)
Overall_Incidence <- HFMD_Penang_Overall %>% mutate(Incidence_1000 = (n/PENDUDUK_L)*1000)
```
# 5. Merge Penang and HFMD Incidence Data
```{r}
HFMD_Penang_Overall_Incidence <- st_join(Penang, Overall_Incidence)
glimpse(HFMD_Penang_Overall_Incidence)
```
# 6.Plot HFMD Incidence in Pulau Pinang (Overall)
```{r}
tm_shape(Penang) + tm_borders(alpha=.6)  + tm_shape(HFMD_Penang_Overall_Incidence) + tm_polygons("Incidence_1000",style = "jenks", title = "Crude Incidence of HFMD per 1000 Population", palette = "Reds", range = c(0, 1000), breaks= 10)+tm_layout( legend.outside = TRUE) 
```
# 7.Plot Yearly Incidence Rate of HFMD in Pulau Pinang
```{r}
tm_shape(Penang) + tm_borders(alpha=.6)  + tm_shape(HFMD_Penang_Overall_Incidence) + tm_polygons("Incidence_1000",style = "jenks", title = "Crude Incidence of HFMD per 1000 Population", palette = "Reds", range = c(0, 1000), breaks= 10) +  tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
```
# 8. Yearly incidence according to district in Pulau Pinang
# Timur Laut
```{r}
tm_shape(Penang[Penang$NAMA_DP == "TIMUR LAUT", ])+ tm_borders(alpha=.6) +
tm_shape(HFMD_Penang_Overall_Incidence[HFMD_Penang_Overall_Incidence$ NAMA_DP.x== "TIMUR LAUT",])+tm_polygons("Incidence_1000",style = "jenks", title = "(Incidence per 1000 Population)", palette = "Reds") + tm_facets("YEAR", ncol = 6, as.layers = TRUE)+tm_layout(title = "INCIDENCE OF HFMD IN TIMUR LAUT",title.size=1)+  tm_layout(legend.outside = TRUE)
```
#  Barat Daya
```{r}
tm_shape(Penang[Penang$NAMA_DP == "BARAT DAYA", ])+ tm_borders(alpha=.6) +
tm_shape(HFMD_Penang_Overall_Incidence[HFMD_Penang_Overall_Incidence$ NAMA_DP.x== "BARAT DAYA",])+tm_polygons("Incidence_1000",style = "jenks", title = "(Incidence per 1000 Population)", palette = "Reds") + tm_facets("YEAR", ncol = 6, as.layers = TRUE)+tm_layout(title = "INCIDENCE OF HFMD IN BARAT DAYA",title.size=1)+  tm_layout(legend.outside = TRUE)
```
#  SP Tengah
```{r}
tm_shape(Penang[Penang$NAMA_DP == "SP.TENGAH", ])+ tm_borders(alpha=.6) +
tm_shape(HFMD_Penang_Overall_Incidence[HFMD_Penang_Overall_Incidence$ NAMA_DP.x== "SP.TENGAH",])+tm_polygons("Incidence_1000",style = "jenks", title = "(Incidence per 1000 Population)", palette = "Reds") + tm_facets("YEAR", ncol = 6, as.layers = TRUE)+tm_layout(title = "INCIDENCE OF HFMD IN SP.PERAI TENGAH",title.size=1)+  tm_layout(legend.outside = TRUE)
```
# SP Utara
```{r}
tm_shape(Penang[Penang$NAMA_DP == "SP.UTARA", ])+ tm_borders(alpha=.6) +
tm_shape(HFMD_Penang_Overall_Incidence[HFMD_Penang_Overall_Incidence$ NAMA_DP.x== "SP.UTARA",])+tm_polygons("Incidence_1000",style = "jenks", title = "(Incidence per 1000 Population)", palette = "Reds") + tm_facets("YEAR", ncol = 6, as.layers = TRUE)+tm_layout(title = "INCIDENCE OF HFMD IN SP.UTARA",title.size=1)+  tm_layout(legend.outside = TRUE)
```
# SP Selatan
```{r}
tm_shape(Penang[Penang$NAMA_DP == "SP.SELATAN", ])+ tm_borders(alpha=.6) +
tm_shape(HFMD_Penang_Overall_Incidence[HFMD_Penang_Overall_Incidence$ NAMA_DP.x== "SP.SELATAN",])+tm_polygons("Incidence_1000",style = "jenks", title = "(Incidence per 1000 Population)", palette = "Reds") + tm_facets("YEAR", ncol = 6, as.layers = TRUE)+tm_layout(title = "INCIDENCE OF HFMD IN SP.SELATAN",title.size=1)+  tm_layout(legend.outside = TRUE)
```
# Density Based Analysis using Kernel Density Estimates (KDE)
# Rescale and load point feature into Penang.owin
#Move coordinate to first column
```{r}
HFMD_overall <- HFMD_overall[, c("geometry", names(HFMD_overall)[-which(names(HFMD_overall) == "geometry")])]
```
#Create owin data for Penang
```{r}
Penang.owin <- as.owin(Penang)
Penang.owin.km <- rescale(Penang.owin, 1000)
```
#Load point feature shapefile into Penang.owin
```{r}
HFMD_overall.ppp  <- as.ppp(HFMD_overall)
marks(HFMD_overall.ppp) <- NULL
HFMD_overall.ppp <- rescale(HFMD_overall.ppp, 1000,"km")
Window(HFMD_overall.ppp) <- Penang.owin.km
```
#Kernel Density Estimates (KDE) Pulau Pinang
```{r}
HFMD_overall.density <- density(HFMD_overall.ppp, sigma=1) 
data.class(HFMD_overall.density)
```
#Rasterize the KDE image using package raster
```{r}
library(raster)
library(viridis)
library(leaflet)
```
```{r}
HFMD_overall.density.raster <- raster(HFMD_overall.density)
```
#Change the extent of the rasterlayer
#1.define intended extend objects using Kelantan polygon
```{r}
xmin= 243501.5 
xmax= 285076.2
ymin= 567201.1 
ymax= 618503.2
```
#2.grab the cols and rows for the raster using @ncols and @nrows
```{r}
HFMD_overall.density.raster @ncols
HFMD_overall.density.raster @nrows
```
#3.create a raster extend class
```{r}
rasExt <- extent(xmin,xmax,ymin,ymax)
rasExt
```
#4.apply the extent to our KDE raster
```{r}
HFMD_overall.density.raster@extent <- rasExt
HFMD_overall.density.raster
```
```{r}
crs(HFMD_overall.density.raster) <- 3168
```
## Plot Intensity of HFMD in Pulau Pinang
```{r}
tm_shape(Penang) +
  tm_borders(alpha=.6) +
  tm_shape(HFMD_overall.density.raster) + 
  tm_raster(style = "quantile", n = 7 , palette = viridis(7, direction = 1), alpha = .4, title = "HFMD Intensity (Cases per KM square)") + 
  tm_layout(legend.outside = TRUE)
```
## KDE intensity by YEAR
#Convert "YEAR" variable to factor with ordered levels
```{r}
HFMD_overall$YEAR <- factor(HFMD_overall$YEAR, 
                             ordered = TRUE)
```
#Plot Case vs YEAR
```{r}
counts <- table(HFMD_overall$YEAR)
barplot(counts, main = "HFMD Infections", xlab = "Year",
        ylab = "Number of Cases", cex.axis = 1, cex.lab = 1, cex.names = 0.7, ylim = c(0, 6500))
text(x = 1:length(counts), y = counts, labels = counts, pos = 3, cex = 0.8)
```
#KDE density for each year
```{r}
Coor_Year <- HFMD_overall[,c(1,3,4)]
Coor_Year  <- na.omit(Coor_Year )
```
#Specified coordinate for male and female
```{r}
Y2017_coor <- subset(Coor_Year, YEAR=='2017') 
Y2018_coor <- subset(Coor_Year, YEAR=='2018') 
Y2019_coor <- subset(Coor_Year, YEAR=='2019') 
Y2020_coor <- subset(Coor_Year, YEAR=='2020') 
Y2021_coor <- subset(Coor_Year, YEAR=='2021') 
Y2022_coor <- subset(Coor_Year, YEAR=='2022') 
```
#Load a point feature shapefile into Penang.owin
```{r}
Y2017_coor.ppp  <- as.ppp(Y2017_coor)
marks(Y2017_coor.ppp ) <- NULL
Y2017_coor.ppp  <- rescale(Y2017_coor.ppp , 1000,"km")
Window(Y2017_coor.ppp ) <- Penang.owin.km
Y2018_coor.ppp  <- as.ppp(Y2018_coor)
marks(Y2018_coor.ppp ) <- NULL
Y2018_coor.ppp  <- rescale(Y2018_coor.ppp , 1000,"km")
Window(Y2018_coor.ppp ) <- Penang.owin.km
Y2019_coor.ppp  <- as.ppp(Y2019_coor)
marks(Y2019_coor.ppp ) <- NULL
Y2019_coor.ppp  <- rescale(Y2019_coor.ppp , 1000,"km")
Window(Y2019_coor.ppp ) <- Penang.owin.km
Y2020_coor.ppp  <- as.ppp(Y2020_coor)
marks(Y2017_coor.ppp ) <- NULL
Y2020_coor.ppp  <- rescale(Y2020_coor.ppp , 1000,"km")
Window(Y2020_coor.ppp ) <- Penang.owin.km
Y2021_coor.ppp  <- as.ppp(Y2021_coor)
marks(Y2021_coor.ppp ) <- NULL
Y2021_coor.ppp  <- rescale(Y2021_coor.ppp , 1000,"km")
Window(Y2021_coor.ppp ) <- Penang.owin.km
Y2022_coor.ppp  <- as.ppp(Y2022_coor)
marks(Y2022_coor.ppp ) <- NULL
Y2022_coor.ppp  <- rescale(Y2022_coor.ppp , 1000,"km")
Window(Y2022_coor.ppp ) <- Penang.owin.km
```
#Create Density function for gender (bandwith 2KM)
```{r}
Y2017.coor.density<- density(Y2017_coor.ppp, sigma=2) 
Y2018.coor.density<- density(Y2018_coor.ppp, sigma=2) 
Y2019.coor.density<- density(Y2019_coor.ppp, sigma=2) 
Y2020.coor.density<- density(Y2020_coor.ppp, sigma=2) 
Y2021.coor.density<- density(Y2021_coor.ppp, sigma=2) 
Y2022.coor.density<- density(Y2022_coor.ppp, sigma=2) 
```
#Plot KDE for Yearly Case
```{r}
plot(Y2017.coor.density, main = "2017", las = 1)
contour(Y2017.coor.density, add = TRUE)
plot(Y2018.coor.density, main = "2018", las = 1)
contour(Y2018.coor.density, add = TRUE)
plot(Y2019.coor.density, main = "2019", las = 1)
contour(Y2019.coor.density, add = TRUE)
plot(Y2020.coor.density, main = "2020", las = 1)
contour(Y2020.coor.density, add = TRUE)
plot(Y2021.coor.density, main = "2021", las = 1)
contour(Y2021.coor.density, add = TRUE)
plot(Y2022.coor.density, main = "2022", las = 1)
contour(Y2022.coor.density, add = TRUE)
```
# KDE intensity by Gender
## Select Gender and Coordinate from data
```{r}
Coor_Gender <- HFMD_overall[,c(1,8)]
Coor_Gender <- na.omit(Coor_Gender)
```
#Specified coordinate for male and female
```{r}
Male.coor <- subset(Coor_Gender, GENDER=='Lelaki') 
Female.coor <- subset(Coor_Gender, GENDER=='Perempuan') 
```
#Load a point feature shapefile into Penang.owin
```{r}
Male.coor.ppp  <- as.ppp(Male.coor)
marks(Male.coor.ppp ) <- NULL
Male.coor.ppp  <- rescale(Male.coor.ppp , 1000,"km")
Window(Male.coor.ppp ) <- Penang.owin.km
Female.coor.ppp  <- as.ppp(Female.coor)
marks(Female.coor.ppp ) <- NULL
Female.coor.ppp  <- rescale(Female.coor.ppp , 1000,"km")
Window(Female.coor.ppp ) <- Penang.owin.km
```
#Create Density function for gender
```{r}
Male.coor.density<- density(Male.coor.ppp , sigma=2) 
Female.coor.density <- density(Female.coor.ppp , sigma=2) 
```
#Plot KDE for gender
```{r}
par(mfrow=c(1,2))
plot(Male.coor.density, main = "Male", las = 1)
contour(Male.coor.density, add = TRUE)
plot(Female.coor.density, main = "Female", las = 1)
contour(Female.coor.density, add = TRUE)
```
#KDE intensity by Age Group
#Categorise age of the cases
```{r}
HFMD_overall ["age_group"] = cut (HFMD_overall$AGE, c (0,5,10,Inf),c("0-4","5-10",">10"), include.lowest = TRUE)
```
#Plot Case vs Age Group
```{r}
counts <- table(HFMD_overall$age_group)
barplot(counts, main = "HFMD Infections", xlab = "Age Group",
        ylab = "Number of Cases", cex.axis = 1, cex.lab = 1, cex.names = 0.7, ylim = c(0, 14000))
text(x = 1:length(counts), y = counts, labels = counts, pos = 3, cex = 0.8)
```
#KDE density for each age group
```{r}
Coor_Age <- HFMD_overall[,c(1,13)]
Coor_Age  <- na.omit(Coor_Age )
```
#Specified coordinate for each age group
```{r}
Coor_Age_1 <- subset(Coor_Age, age_group =='0-4') 
Coor_Age_2 <- subset(Coor_Age, age_group =='5-10')  
Coor_Age_3 <- subset(Coor_Age, age_group =='>10') 
```
#Load a point feature shapefile into Penang.owin
```{r}
Coor_Age_1.ppp  <- as.ppp(Coor_Age_1)
marks(Coor_Age_1.ppp ) <- NULL
Coor_Age_1.ppp  <- rescale(Coor_Age_1.ppp , 1000,"km")
Window(Coor_Age_1.ppp ) <- Penang.owin.km
Coor_Age_2.ppp  <- as.ppp(Coor_Age_2)
marks(Coor_Age_2.ppp ) <- NULL
Coor_Age_2.ppp  <- rescale(Coor_Age_2.ppp , 1000,"km")
Window(Coor_Age_2.ppp ) <- Penang.owin.km
Coor_Age_3.ppp  <- as.ppp(Coor_Age_3)
marks(Coor_Age_3.ppp ) <- NULL
Coor_Age_3.ppp  <- rescale(Coor_Age_3.ppp , 1000,"km")
Window(Coor_Age_3.ppp ) <- Penang.owin.km
```
#Create Density function for each age group
```{r}
Coor_Age_1.density<- density(Coor_Age_1.ppp, sigma=2) 
Coor_Age_2.density<- density(Coor_Age_2.ppp, sigma=2) 
Coor_Age_3.density<- density(Coor_Age_3.ppp, sigma=2) 
```
#Plot KDE for each age group
```{r}
par(mfrow=c(1,3))
plot(Coor_Age_1.density, main = "0-4", las = 1)
contour(Coor_Age_1.density, add = TRUE)
plot(Coor_Age_2.density, main = "5-10", las = 1)
contour(Coor_Age_2.density, add = TRUE)
plot(Coor_Age_3.density, main = ">10", las = 1)
contour(Coor_Age_3.density, add = TRUE)
```


