---
title: "Spatial Autocorrelation Analysis"
author: "Dr Afiq"
date: "2023-03-22"
output: html_document
---
```{r}
library(spdep)
library(sf)
library (readxl)
library (here)
library (ggplot2)
```
# Load Data
```{r}
df_17 <- read_excel( here("Spatial Autocorrelation",
                          "count_HFMD_Mukim_17_Cases.xlsx"))
df_18 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2018_Mukim.xlsx"))
df_19 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2019_Mukim.xlsx"))
df_20 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2020_Mukim.xlsx"))
df_21 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2021_Mukim.xlsx"))
df_22 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2022_Mukim.xlsx"))
Penang <- st_read(here("Penang", 
                  "mukim2000_penang1.shp"))
```
```{r}
st_crs(Penang)<-3168
```

# Merge with Penang Polygon
```{r}
HFMD_Poly_17 <- merge(Penang, df_17, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_18 <- merge(Penang, df_18, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_19 <- merge(Penang, df_19, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_20 <- merge(Penang, df_20, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_21 <- merge(Penang, df_21, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_22 <- merge(Penang, df_22, by = c("NAMA_DP", "NAMMUK"))
```
# Set Neighbouring Polygon for autocorrelation analysis (using QUEEN approach)
```{r}
nb_17 <- poly2nb(HFMD_Poly_17, queen = TRUE)
nb_18 <- poly2nb(HFMD_Poly_18, queen = TRUE)
nb_19 <- poly2nb(HFMD_Poly_19, queen = TRUE)
nb_20 <- poly2nb(HFMD_Poly_20, queen = TRUE)
nb_21 <- poly2nb(HFMD_Poly_21, queen = TRUE)
nb_22 <- poly2nb(HFMD_Poly_22, queen = TRUE)
```
# view Neighbour for (1st Mukim)
```{r}
nb_17[[1]]
nb_18 [[1]]
```
#Assign Weight for Mukim that contribute to Case in particular mukim
```{r}
lw_17 <- nb2listw(nb_17, style = "W" , zero.policy = TRUE)
lw_18 <- nb2listw(nb_18, style = "W" , zero.policy = TRUE)
lw_19 <- nb2listw(nb_19, style = "W" , zero.policy = TRUE)
lw_20 <- nb2listw(nb_20, style = "W" , zero.policy = TRUE)
lw_21 <- nb2listw(nb_21, style = "W" , zero.policy = TRUE)
lw_22 <- nb2listw(nb_22, style = "W" , zero.policy = TRUE)
```
# Create Lag Function for Moran Test
```{r}
HFMD_17_lag <- lag.listw(lw_17, HFMD_Poly_17$n)
HFMD_18_lag <- lag.listw(lw_18, HFMD_Poly_18$n)
HFMD_19_lag <- lag.listw(lw_19, HFMD_Poly_19$n)
HFMD_20_lag <- lag.listw(lw_20, HFMD_Poly_20$n)
HFMD_21_lag <- lag.listw(lw_21, HFMD_Poly_21$n)
HFMD_22_lag <- lag.listw(lw_22, HFMD_Poly_22$n)
```
# Computing Global Moran I for cases in 2017
```{r}
moran.test(HFMD_Poly_17$n, lw_17)
moran.test(HFMD_Poly_18$n, lw_18)
moran.test(HFMD_Poly_19$n, lw_19)
moran.test(HFMD_Poly_20$n, lw_20)
moran.test(HFMD_Poly_21$n, lw_21)
moran.test(HFMD_Poly_22$n, lw_22)
```
# Computing Local Moran Analysis
```{r}
Local_Moran_17 <- localmoran(HFMD_Poly_17$n,lw_17)
Local_Moran_18 <- localmoran(HFMD_Poly_18$n,lw_18)
Local_Moran_19 <- localmoran(HFMD_Poly_19$n,lw_19)
Local_Moran_20 <- localmoran(HFMD_Poly_20$n,lw_20)
Local_Moran_21 <- localmoran(HFMD_Poly_21$n,lw_21)
Local_Moran_22 <- localmoran(HFMD_Poly_22$n,lw_22)
```
# Join results to sf data
```{r}
HFMD_Poly_17 <- cbind(HFMD_Poly_17, Local_Moran_17)
HFMD_Poly_18 <- cbind(HFMD_Poly_18, Local_Moran_18)
HFMD_Poly_19 <- cbind(HFMD_Poly_19, Local_Moran_19)
HFMD_Poly_20 <- cbind(HFMD_Poly_20, Local_Moran_20)
HFMD_Poly_21 <- cbind(HFMD_Poly_21, Local_Moran_21)
HFMD_Poly_22 <- cbind(HFMD_Poly_22, Local_Moran_22)
```
#Plot Local Moran Statistic 17
```{r}
ggplot(data = HFMD_Poly_17) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2017",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_18) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2018",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_19) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2019",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_20) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2020",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_21) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2021",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_22) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2022",
       subtitle="Pulau Pinang")
```

## LISA (Local Indicator od Spatial Association)

#Center the variable of interest around its mean
```{r}
m_Q_17 <- HFMD_Poly_17$n - mean(HFMD_Poly_17$n)
m_Q_18 <- HFMD_Poly_18$n - mean(HFMD_Poly_18$n)
m_Q_19 <- HFMD_Poly_19$n - mean(HFMD_Poly_19$n)
m_Q_20 <- HFMD_Poly_20$n - mean(HFMD_Poly_20$n)
m_Q_21 <- HFMD_Poly_21$n - mean(HFMD_Poly_21$n)
m_Q_22 <- HFMD_Poly_22$n - mean(HFMD_Poly_22$n)
```
# Center the local Moran around the mean
```{r}
m_Q_Local_17 <- Local_Moran_17[,1]- mean(Local_Moran_17[,1])
m_Q_Local_18 <- Local_Moran_18[,1]- mean(Local_Moran_18[,1])
m_Q_Local_19 <- Local_Moran_19[,1]- mean(Local_Moran_19[,1])
m_Q_Local_20 <- Local_Moran_20[,1]- mean(Local_Moran_20[,1])
m_Q_Local_21 <- Local_Moran_21[,1]- mean(Local_Moran_21[,1])
m_Q_Local_22 <- Local_Moran_22[,1]- mean(Local_Moran_22[,1])
```
#significance threshold
```{r}
signif <- 0.05
```
#builds a data quadrant
```{r}
quadrant_17 <- vector(mode="numeric",length = nrow(Local_Moran_17))
quadrant_18 <- vector(mode="numeric",length = nrow(Local_Moran_18))
quadrant_19 <- vector(mode="numeric",length = nrow(Local_Moran_19))
quadrant_20 <- vector(mode="numeric",length = nrow(Local_Moran_20))
quadrant_21 <- vector(mode="numeric",length = nrow(Local_Moran_21))
quadrant_22 <- vector(mode="numeric",length = nrow(Local_Moran_22))
```
```{r}
quadrant_17 [m_Q_17>0 & m_Q_Local_17>0] <- 4
quadrant_17 [m_Q_17<0 & m_Q_Local_17<0] <- 1
quadrant_17 [m_Q_17<0 & m_Q_Local_17>0] <- 2
quadrant_17 [m_Q_17>0 & m_Q_Local_17<0] <- 3
quadrant_17 [Local_Moran_17[,5]] <- 0
quadrant_18 [m_Q_18>0 & m_Q_Local_18>0] <- 4
quadrant_18 [m_Q_18<0 & m_Q_Local_18<0] <- 1
quadrant_18 [m_Q_18<0 & m_Q_Local_18>0] <- 2
quadrant_18 [m_Q_18>0 & m_Q_Local_18<0] <- 3
quadrant_18 [Local_Moran_18[,5]] <- 0
quadrant_19 [m_Q_19>0 & m_Q_Local_19>0] <- 4
quadrant_19 [m_Q_19<0 & m_Q_Local_19<0] <- 1
quadrant_19 [m_Q_19<0 & m_Q_Local_19>0] <- 2
quadrant_19 [m_Q_19>0 & m_Q_Local_19<0] <- 3
quadrant_19 [Local_Moran_19[,5]] <- 0
quadrant_20 [m_Q_20>0 & m_Q_Local_20>0] <- 4
quadrant_20 [m_Q_20<0 & m_Q_Local_20<0] <- 1
quadrant_20 [m_Q_20<0 & m_Q_Local_20>0] <- 2
quadrant_20 [m_Q_20>0 & m_Q_Local_20<0] <- 3
quadrant_20 [Local_Moran_20[,5]] <- 0
quadrant_21 [m_Q_21>0 & m_Q_Local_21>0] <- 4
quadrant_21 [m_Q_21<0 & m_Q_Local_21<0] <- 1
quadrant_21 [m_Q_21<0 & m_Q_Local_21>0] <- 2
quadrant_21 [m_Q_21>0 & m_Q_Local_21<0] <- 3
quadrant_21[Local_Moran_21[,5]] <- 0
quadrant_22 [m_Q_22>0 & m_Q_Local_22>0] <- 4
quadrant_22 [m_Q_22<0 & m_Q_Local_22<0] <- 1
quadrant_22 [m_Q_22<0 & m_Q_Local_22>0] <- 2
quadrant_22 [m_Q_22>0 & m_Q_Local_22<0] <- 3
quadrant_22[Local_Moran_22[,5]] <- 0
```
# plot LISA Map
```{r}
brks <- c(0,1,2,3,4)
colors <- c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red")
plot(HFMD_Poly_17[,5],main = "LISA Map 2017", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_18[,6],main = "LISA Map 2018", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_19[,6],main = "LISA Map 2019", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_20[,6],main = "LISA Map 2020", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_21[,6],main = "LISA Map 2021", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_22[,6],main = "LISA Map 2022", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
```



