---
title: "Spatial Autocorrelation Analysis"
author: "Dr Afiq"
date: "2023-03-22"
output: html_document
---
```{r}
library(spdep)
library(sf)
library (readxl)
library (here)
library (ggplot2)
library (tmap)
library(leaflet)
library (RColorBrewer)
```
# Load Data
```{r}
df_17 <- read_excel( here("Spatial Autocorrelation",
                          "count_HFMD_Mukim_17_Cases.xlsx"))
df_18 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2018_Mukim.xlsx"))
df_19 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2019_Mukim.xlsx"))
df_20 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2020_Mukim.xlsx"))
df_21 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2021_Mukim.xlsx"))
df_22 <- read_excel( here("Spatial Autocorrelation",
                          "HFMD_2022_Mukim.xlsx"))
Penang <- st_read(here("Penang", 
                  "mukim2000_penang1.shp"))
```
```{r}
st_crs(Penang)<-3168
```

# Merge with Penang Polygon
```{r}
HFMD_Poly_17 <- merge(Penang, df_17, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_18 <- merge(Penang, df_18, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_19 <- merge(Penang, df_19, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_20 <- merge(Penang, df_20, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_21 <- merge(Penang, df_21, by = c("NAMA_DP", "NAMMUK"))
HFMD_Poly_22 <- merge(Penang, df_22, by = c("NAMA_DP", "NAMMUK"))
```
# Set Neighbouring Polygon for autocorrelation analysis (using QUEEN approach)
```{r}
nb_17 <- poly2nb(HFMD_Poly_17, queen = TRUE)
nb_18 <- poly2nb(HFMD_Poly_18, queen = TRUE)
nb_19 <- poly2nb(HFMD_Poly_19, queen = TRUE)
nb_20 <- poly2nb(HFMD_Poly_20, queen = TRUE)
nb_21 <- poly2nb(HFMD_Poly_21, queen = TRUE)
nb_22 <- poly2nb(HFMD_Poly_22, queen = TRUE)
```
# view Neighbour for (1st Mukim)
```{r}
nb_17[[1]]
nb_18 [[1]]
```
#Assign Weight for Mukim that contribute to Case in particular mukim
```{r}
lw_17 <- nb2listw(nb_17, style = "W" , zero.policy = TRUE)
lw_18 <- nb2listw(nb_18, style = "W" , zero.policy = TRUE)
lw_19 <- nb2listw(nb_19, style = "W" , zero.policy = TRUE)
lw_20 <- nb2listw(nb_20, style = "W" , zero.policy = TRUE)
lw_21 <- nb2listw(nb_21, style = "W" , zero.policy = TRUE)
lw_22 <- nb2listw(nb_22, style = "W" , zero.policy = TRUE)
```
# Create Lag Function for Moran Test
```{r}
HFMD_17_lag <- lag.listw(lw_17, HFMD_Poly_17$n)
HFMD_18_lag <- lag.listw(lw_18, HFMD_Poly_18$n)
HFMD_19_lag <- lag.listw(lw_19, HFMD_Poly_19$n)
HFMD_20_lag <- lag.listw(lw_20, HFMD_Poly_20$n)
HFMD_21_lag <- lag.listw(lw_21, HFMD_Poly_21$n)
HFMD_22_lag <- lag.listw(lw_22, HFMD_Poly_22$n)
```
# Computing Global Moran I for cases in 2017
```{r}
moran.test(HFMD_Poly_17$n, lw_17)
moran.test(HFMD_Poly_18$n, lw_18)
moran.test(HFMD_Poly_19$n, lw_19)
moran.test(HFMD_Poly_20$n, lw_20)
moran.test(HFMD_Poly_21$n, lw_21)
moran.test(HFMD_Poly_22$n, lw_22)
```
# Computing Local Moran Analysis
```{r}
Local_Moran_17 <- localmoran(HFMD_Poly_17$n,lw_17)
Local_Moran_18 <- localmoran(HFMD_Poly_18$n,lw_18)
Local_Moran_19 <- localmoran(HFMD_Poly_19$n,lw_19)
Local_Moran_20 <- localmoran(HFMD_Poly_20$n,lw_20)
Local_Moran_21 <- localmoran(HFMD_Poly_21$n,lw_21)
Local_Moran_22 <- localmoran(HFMD_Poly_22$n,lw_22)
```
# Join results to sf data
```{r}
HFMD_Poly_17 <- cbind(HFMD_Poly_17, Local_Moran_17)
HFMD_Poly_18 <- cbind(HFMD_Poly_18, Local_Moran_18)
HFMD_Poly_19 <- cbind(HFMD_Poly_19, Local_Moran_19)
HFMD_Poly_20 <- cbind(HFMD_Poly_20, Local_Moran_20)
HFMD_Poly_21 <- cbind(HFMD_Poly_21, Local_Moran_21)
HFMD_Poly_22 <- cbind(HFMD_Poly_22, Local_Moran_22)
```

#Plot Local Moran Statistic 
```{r}
ggplot(data = HFMD_Poly_17) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2017",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_18) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2018",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_19) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2019",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_20) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2020",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_21) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2021",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_22) +
  geom_sf(aes(fill=Ii)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Local Moran's I") +
  labs(title="Local Moran's I statistic for HFMD in 2022",
       subtitle="Pulau Pinang")
```
# Plot Local Moran Using TMAP
```{r name-of-chunk, fig.width=9, fig.height=5}
LM_17<-tm_shape(Penang) + 
  tm_fill(col="#f7f7f7")+ tm_style("natural")+ 
tm_shape(HFMD_Poly_17) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") + tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2017",title.size = 1) 
LM_18<-tm_shape(Penang) + 
  tm_fill(col="#f7f7f7")+ tm_style("natural")+
tm_shape(HFMD_Poly_18) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") +tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2018",title.size = 1)
LM_19<-tm_shape(Penang) + 
 tm_fill(col="#f7f7f7")+ tm_style("natural")+
tm_shape(HFMD_Poly_19) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") +tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2019",title.size = 1) 
LM_20<-tm_shape(Penang) + 
  tm_polygons(col="#f7f7f7")+ tm_style("natural")+
tm_shape(HFMD_Poly_20) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") +tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2020",title.size = 1) 
LM_21<-tm_shape(Penang) + 
  tm_polygons(col="#f7f7f7")+ tm_style("natural")+
tm_shape(HFMD_Poly_21) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") +tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2021",title.size = 1) 
LM_22<-tm_shape(Penang) + 
 tm_fill(col="#f7f7f7")+ tm_style("natural")+
tm_shape(HFMD_Poly_22) +
  tm_polygons(col = "Ii", title = "Local Moran's I", style = "jenks",palette = "Reds") +tm_scale_bar(position = c("right", "top")) +
  tm_layout(title = "2022",title.size = 1) 
```

```{r name-of-chunk, fig.width=11, fig.height=9}
tmap_arrange(LM_17,LM_18,LM_19,LM_20,LM_21,LM_22, ncol = 3) 
```

```{r}
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_17) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = "2017",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_18) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = " 2018",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_19) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = "2019",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_20) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = " 2020",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_21) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = " 2021",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_22) +
  tm_polygons(col = "Pr.z....E.Ii..", title = "Significance level", breaks = c(-Inf,0.001,0.01,0.05,Inf),palette = "-YlOrRd") +
  tm_layout(main.title = "2022",title.size = 1.5) +
tm_compass(type = "4star", size = 2, position = c("left", "top"))
```

## LISA (Local Indicator od Spatial Association)

#Center the variable of interest around its mean
```{r}
m_Q_17 <- HFMD_Poly_17$n - mean(HFMD_Poly_17$n)
m_Q_18 <- HFMD_Poly_18$n - mean(HFMD_Poly_18$n)
m_Q_19 <- HFMD_Poly_19$n - mean(HFMD_Poly_19$n)
m_Q_20 <- HFMD_Poly_20$n - mean(HFMD_Poly_20$n)
m_Q_21 <- HFMD_Poly_21$n - mean(HFMD_Poly_21$n)
m_Q_22 <- HFMD_Poly_22$n - mean(HFMD_Poly_22$n)
```
# Center the local Moran around the mean
```{r}
m_Q_Local_17 <- Local_Moran_17[,1]- mean(Local_Moran_17[,1])
m_Q_Local_18 <- Local_Moran_18[,1]- mean(Local_Moran_18[,1])
m_Q_Local_19 <- Local_Moran_19[,1]- mean(Local_Moran_19[,1])
m_Q_Local_20 <- Local_Moran_20[,1]- mean(Local_Moran_20[,1])
m_Q_Local_21 <- Local_Moran_21[,1]- mean(Local_Moran_21[,1])
m_Q_Local_22 <- Local_Moran_22[,1]- mean(Local_Moran_22[,1])
```
#significance threshold
```{r}
signif <- 0.05
```
#builds a data quadrant
```{r}
quadrant_17 <- vector(mode="numeric",length = nrow(Local_Moran_17))
quadrant_18 <- vector(mode="numeric",length = nrow(Local_Moran_18))
quadrant_19 <- vector(mode="numeric",length = nrow(Local_Moran_19))
quadrant_20 <- vector(mode="numeric",length = nrow(Local_Moran_20))
quadrant_21 <- vector(mode="numeric",length = nrow(Local_Moran_21))
quadrant_22 <- vector(mode="numeric",length = nrow(Local_Moran_22))
```
```{r}
quadrant_17 [m_Q_17>0 & m_Q_Local_17>0] <- 4
quadrant_17 [m_Q_17<0 & m_Q_Local_17<0] <- 1
quadrant_17 [m_Q_17<0 & m_Q_Local_17>0] <- 2
quadrant_17 [m_Q_17>0 & m_Q_Local_17<0] <- 3
quadrant_17 [Local_Moran_17[,5]] <- 0
quadrant_18 [m_Q_18>0 & m_Q_Local_18>0] <- 4
quadrant_18 [m_Q_18<0 & m_Q_Local_18<0] <- 1
quadrant_18 [m_Q_18<0 & m_Q_Local_18>0] <- 2
quadrant_18 [m_Q_18>0 & m_Q_Local_18<0] <- 3
quadrant_18 [Local_Moran_18[,5]] <- 0
quadrant_19 [m_Q_19>0 & m_Q_Local_19>0] <- 4
quadrant_19 [m_Q_19<0 & m_Q_Local_19<0] <- 1
quadrant_19 [m_Q_19<0 & m_Q_Local_19>0] <- 2
quadrant_19 [m_Q_19>0 & m_Q_Local_19<0] <- 3
quadrant_19 [Local_Moran_19[,5]] <- 0
quadrant_20 [m_Q_20>0 & m_Q_Local_20>0] <- 4
quadrant_20 [m_Q_20<0 & m_Q_Local_20<0] <- 1
quadrant_20 [m_Q_20<0 & m_Q_Local_20>0] <- 2
quadrant_20 [m_Q_20>0 & m_Q_Local_20<0] <- 3
quadrant_20 [Local_Moran_20[,5]] <- 0
quadrant_21 [m_Q_21>0 & m_Q_Local_21>0] <- 4
quadrant_21 [m_Q_21<0 & m_Q_Local_21<0] <- 1
quadrant_21 [m_Q_21<0 & m_Q_Local_21>0] <- 2
quadrant_21 [m_Q_21>0 & m_Q_Local_21<0] <- 3
quadrant_21[Local_Moran_21[,5]] <- 0
quadrant_22 [m_Q_22>0 & m_Q_Local_22>0] <- 4
quadrant_22 [m_Q_22<0 & m_Q_Local_22<0] <- 1
quadrant_22 [m_Q_22<0 & m_Q_Local_22>0] <- 2
quadrant_22 [m_Q_22>0 & m_Q_Local_22<0] <- 3
quadrant_22[Local_Moran_22[,5]] <- 0
```
# plot LISA Map
```{r}
brks <- c(0,1,2,3,4)
colors <- c("grey", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red")
plot(HFMD_Poly_17[,5],main = "LISA Map 2017", las = 1, border="lightgray",col=colors[findInterval(quadrant_17,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_18[,6],main = "LISA Map 2018", las = 1, border="lightgray",col=colors[findInterval(quadrant_18,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_19[,6],main = "LISA Map 2019", las = 1, border="lightgray",col=colors[findInterval(quadrant_19,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_20[,6],main = "LISA Map 2020", las = 1, border="lightgray",col=colors[findInterval(quadrant_20,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_21[,6],main = "LISA Map 2021", las = 1, border="lightgray",col=colors[findInterval(quadrant_21,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
plot(HFMD_Poly_22[,6],main = "LISA Map 2022", las = 1, border="black",col=colors[findInterval(quadrant_22,brks,all.inside = FALSE)])
legend("bottomright",legend = c("insignificant","low-low","low-high","high-low","high-high"),
       fill=colors,bty="n")
```
```{r}
#bind quadrant and Poly data
LISA_17 <- cbind(HFMD_Poly_17, quadrant_17)
LISA_18 <- cbind(HFMD_Poly_18, quadrant_18)
LISA_19 <- cbind(HFMD_Poly_19, quadrant_19)
LISA_20 <- cbind(HFMD_Poly_20, quadrant_20)
LISA_21 <- cbind(HFMD_Poly_21, quadrant_21)
LISA_22 <- cbind(HFMD_Poly_22, quadrant_22)
```

# Plot LISA using TMAP
```{r}
LISA_Map_17 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") + 
tm_shape(LISA_17) + tm_polygons(col = "quadrant_17", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2017",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
LISA_Map_18 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") + 
tm_shape(LISA_18) + tm_polygons(col = "quadrant_18", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2018",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
LISA_Map_19 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") + 
tm_shape(LISA_19) + tm_polygons(col = "quadrant_19", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2019",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
LISA_Map_20 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") +
tm_shape(LISA_20) + tm_polygons(col = "quadrant_20", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2020",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
LISA_Map_21 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") + 
tm_shape(LISA_21) + tm_polygons(col = "quadrant_21", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2021",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
LISA_Map_22 <- tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+ tm_style("natural") +
tm_shape(LISA_22) + tm_polygons(col = "quadrant_22", breaks = c(0,1,2,3,4,5), palette= 
                              c("white", "blue",rgb(0,0,1,alpha=0.4),rgb(1,0,0,alpha = 0.4),"red"),
                                labels = c("Non Significant","Low-Low","Low-High","High-Low", "High-High"), title="") +
 tm_scale_bar(position = c("left", "top")) +
  tm_layout(title = "2022",title.size = 1) + tm_compass(type = "4star", size = 2, position = c("left", "top"))
```
```{r name-of-chunk, fig.width=11, fig.height=9}
tmap_arrange(LISA_Map_17,LISA_Map_18,LISA_Map_19,LISA_Map_20,LISA_Map_21,LISA_Map_22 , ncol = 3) 
```
# Getis - Ord Gi* Statistic 

```{r}
# Perform Local G analysis
getis_ord_17 <- spdep ::localG(
  HFMD_Poly_17$n,lw_17)
getis_ord_18 <- spdep ::localG(
  HFMD_Poly_18$n,lw_18)
getis_ord_19 <- spdep ::localG(
  HFMD_Poly_19$n,lw_19)
getis_ord_20 <- spdep ::localG(
  HFMD_Poly_20$n,lw_20)
getis_ord_21 <- spdep ::localG(
  HFMD_Poly_21$n,lw_21)
getis_ord_22 <- spdep ::localG(
  HFMD_Poly_22$n,lw_22)
```

```{r}
# join the result with SF data
HFMD_Poly_17$getis_ord_17 <- getis_ord_17 
HFMD_Poly_18$getis_ord_18 <- getis_ord_18 
HFMD_Poly_19$getis_ord_19 <- getis_ord_19 
HFMD_Poly_20$getis_ord_20 <- getis_ord_20 
HFMD_Poly_21$getis_ord_21 <- getis_ord_21 
HFMD_Poly_22$getis_ord_22 <- getis_ord_22 
```
```{r}
ggplot(data = HFMD_Poly_17) +
  geom_sf(aes(fill=getis_ord_17)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2017",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_18) +
  geom_sf(aes(fill=getis_ord_18)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2018",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_19) +
  geom_sf(aes(fill=getis_ord_19)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2019",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_20) +
  geom_sf(aes(fill=getis_ord_20)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2020",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_21) +
  geom_sf(aes(fill=getis_ord_21)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2021",
       subtitle="Pulau Pinang")
ggplot(data = HFMD_Poly_22) +
  geom_sf(aes(fill=getis_ord_22)) +
 theme_bw() +
  scale_fill_gradient2(low="#2c7bb6", mid="#ffffbf", high="#d7191c",
                       name="Gi*") +
  labs(title="Getis - Ord Gi* Statistic for HFMD in 2022",
       subtitle="Pulau Pinang")
```
# Population and Cases
```{r}
#plot penang Population
tm_shape(Penang) + tm_borders(alpha=.6)+tm_shape(HFMD_Penang_Overall_Incidence)+ tm_polygons("PENDUDUK_L",title = "Population",title.size = 0.5) + tm_layout(main.title = "Population") +tm_layout(legend.outside = TRUE) +tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
tm_shape(Penang) + tm_borders(alpha=.6) + tm_shape(HFMD_Penang_Overall_Incidence) + tm_polygons("n",title = "HFMD Cases",title.size = 0.5) + tm_layout(main.title = "HFMD Cases") +tm_layout(legend.outside = TRUE) +tm_facets("YEAR", nrow = 2, ncol = 3, as.layers = TRUE)
```
```{r}
#using tmap
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_17) +
  tm_polygons(col = "getis_ord_17", title = "Gi Statistic",
             midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd"))+
  tm_layout(main.title = "2017",title.size = 1) +
  tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_18) +
  tm_polygons(col = "getis_ord_18", title = "Gi Statistic",midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd")) +
  tm_layout(main.title = "2018",title.size = 1) +
  tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_19) +
  tm_polygons(col = "getis_ord_19", title = "Gi Statistic",midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd")) +
  tm_layout(main.title = "2019",title.size = 1) +
  tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_20) +
  tm_polygons(col = "getis_ord_20", title = "Gi Statistic",midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd")) +
  tm_layout(main.title = "2020",title.size = 1) +
 tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_21) +
  tm_polygons(col = "getis_ord_21", title = "Gi Statistic",midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd")) +
 tm_layout(main.title = "2021",title.size = 1) +
  tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_22) +
  tm_polygons(col = "getis_ord_22", title = "Gi Statistic",midpoint = NA,style = "jenks",palette = brewer.pal(5, "YlOrRd")) +
 tm_layout(main.title = "2022",title.size = 1) +
 tm_compass(type = "4star", size = 2, position = c("left", "top"))
```
```{r}
tm_shape(Penang) + 
  tm_polygons(col = "#f7f7f7")+
tm_shape(HFMD_Poly_17) +
  tm_polygons(col = "getis_ord_17", title = "Gi Statistic",midpoint = NA,style = "cont",n = 4,palette = rev(colorRampPalette(c("red", "lightblue"))(100))) +
 tm_layout(main.title = "2017",title.size = 1) +
 tm_compass(type = "4star", size = 2, position = c("left", "top"))
tm_shape(HFMD_Poly_22) +
  tm_polygons(col = "getis_ord_22", title = "Gi Statistic",midpoint = NA,style = "cont",n = 4,palette = rev(colorRampPalette(c("red", "lightblue"))(100))) +
 tm_layout(main.title = "2022",title.size = 1) +
 tm_compass(type = "4star", size = 2, position = c("left", "top"))
```








